---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.path = "man/figures/"
)
```
# jutilities

This a collection of convenience functions for data handling, plotting, and analysis within the `tidyverse` syntax. Most of these functions are wrappers for tasks I perform regularly (e. g., describe variables, add mean indices) and wanted to integrate them into `%>%` pipes with just one line.

This is all very much work in progress.

## Installation

Install from GitHub:

``` r
install.packages("devtools")
devtools::install_github("joon-e/jutilities")
```

## Instructions

`jutilities` are to be used with the `tidyverse`. Most functions will return a tibble and can thus be easily integrated into `%>%` pipes.

```{r message=F, warning=F}
library(tidyverse)
library(jutilities)
```

### Describe continous variables

`describe()` computes several measures of central tendency and variability for all specified variables:

```{r}
diamonds %>% 
  describe(x, y, z)
```

If no variables are specified, all numeric variables are described:

```{r}
diamonds %>% 
  describe()
```

`describe_groups()` outputs the same statistics for one variable, grouped by one or more grouping variables.

```{r}
diamonds %>% 
  describe_groups(price, cut, color)
```
 
### Describe categorical variables

`cat_ftable()` outputs a frequency table including relative, valid, and cumulative frequencies for one categorical variable.

```{r}
diamonds %>% 
  cat_ftable(cut)
```

`cat_xtable` outputs contigency tables for one column variable and one or more row variables:

```{r}
diamonds %>% 
  cat_xtable(cut, color, clarity)
```

Setting the argument `percentages = TRUE` changes the output to relative frequencies:

```{r}
diamonds %>% 
  cat_xtable(cut, color, percentages = TRUE)
```

A Chi² test can be optionally computed by setting the argument `chisq = TRUE`. Test results will be displayed in a console message:

```{r}
diamonds %>% 
  cat_xtable(cut, color, chisq = TRUE)
```

### Modify data

`add_index()` adds a rowwise mean index columns of the specified variables to the dataset. The second argument (first argument if used in a pipe) should be the name of the index column:

```{r}
diamonds %>%
  add_index(meanxyz, x, y, z)
```

Set the argument `type = "sum"` to create a sum index instead:

```{r}
diamonds %>% 
  add_index(sumxyz, x, y, z, type = "sum")
```

`add_label` adds a text label column of a numeric variable to the dataset, for example to be used as labels in plots. 

```{r}
diamonds %>% 
  add_label(labelz, z)
```

By default, `add_label` rounds to two decimal places. You can change this by setting the `decimal.places` argument:

```{r}
diamonds %>% 
  add_label(labelz, z, decimal.places = 0)
```

### Analyze data

#### Correlations

`correlate()` computes correlations coefficients and p-values for all combinations of the specified variables. By default, Pearson correlation coefficients are computed. Set the argument `type = "spearman"` to compute Spearman ranked correlations coefficicents instead.

```{r}
diamonds %>% 
  correlate(x, y, z)
```

#### T-tests

`ttest` computes t-Tests for one grouping variable and one or more test variables. Output statistics include descriptives, t-values, degrees of freedom, p-values, and Cohen's d:

```{r}
diamonds %>% 
  filter(cut == "Ideal" | cut == "Fair") %>% 
  droplevels() %>% 
  ttest(cut, price, x, y, z)
```

